Preventing runtime errors of Redis at compile time

### 2015-11-18 paper.tex
6. Imposing constraints

@@

\section{Denoting containers}
Most Redis commands only work with a certain type of these containers. To
annotate what container a key is associated with, we introduce these container
types.

@@

\subsection{Extending container types}
We rename container types and extend it with an extra type arguement,
to indicate what kind of data it's holding.

@@

\subsection{Imposing constraints}

To rule out programs with undesired properties, certain constraints must be
imposed, on what arguments they can take, and what preconditions they must hold.
These constraints are best encoded in types.

\subsection{Redis datatypes}

Redis supports many different datatypes, and most commands only work with
certain types of them. For example: \mintinline{haskell}{SET} with strings or
\mintinline{haskell}{SADD} with sets.

\section{Discussions}
\section{Conclusions and Related Work}

>>>>

% for clickable footnotes
\usepackage{hyperref}

@@

\subsection{Denoting containers}
Most Redis commands only work with a certain type of these containers. To
annotate what container a key is associated with, we introduce these container
types.

@@

\subsection{Extending container types}
We rename container types and extend it with an extra type arguement,
 to indicate what kind of encoded value it's holding.

@@

\section{Imposing constraints}

To rule out programs with undesired properties, certain constraints must be
imposed, on what arguments they can take, or what preconditions they must hold.

Consider the following example: \mintinline{text}{LLEN} retrurns the length of
the list associated with a key, else raises a type error.

\begin{minted}[xleftmargin=1em]{text}
redis> LPUSH some-list bar
(integer) 1
redis> LLEN some-list
(integer) 1
redis> SET some-string foo
OK
redis> LLEN some-string
(error) WRONGTYPE Operation against a key
holding the wrong kind of value
\end{minted}

Such constraint could be expressed in types with
\emph{equality constraints}\cite{typeeq}.

\begin{minted}[xleftmargin=1em]{haskell}
llen :: (KnownSymbol s, Get xs s ~ Just (ListOf x))
        => Proxy s
        -> Edis xs xs (Either Reply Integer)
llen key = Edis $ Hedis.llen (symbolVal key)
\end{minted}

Where \mintinline{haskell}{(~)} denotes that \mintinline{haskell}{Get xs s}
and \mintinline{haskell}{Just (ListOf x)} needs to be the same.

The semantics of \mintinline{text}{LLEN} defined above is actually not
complete. \mintinline{text}{LLEN} also accepts keys that do not exist and
 returns 0. We will need \emph{}

\begin{minted}[xleftmargin=1em]{text}
redis> LLEN nonexistent
(integer) 0
\end{minted}

In other words, we require that the key to be associated with a list,
 \textbf{or} does not exist at all.

\subsection{Expressing constraint disjunctions}

Unfortunately, expressing disjunctions in constraints is much more difficult
than expressing conjunctions, since the latter could be easily done by placing
 constraints in a tuple (at the left side of \mintinline{haskell}{=>}).

But there are at least three ways to express type-level constraints
\cite{singletons}. Luckily we could express constraint disjunctions with type
 families in a modular way.

The semantics we want could be expressed informally like this:

\mintinline{haskell}{Get xs s} $\equiv$ \mintinline{haskell}{Just (ListOf x)}
$\vee$ $\neg$ \mintinline{haskell}{(Member xs s)}

We could achieve this simply by translating the semantics we want to the
domain of Booleans, with type-level boolean functions such as
\mintinline{haskell}{(&&)},
\mintinline{haskell}{(||)}, \mintinline{haskell}{Not},
\mintinline{haskell}{(==)}, etc.\footnotemark To avoid

\footnotetext{Available in \mintinline{text}{Data.Type.Bool} and
 \mintinline{text}{Data.Type.Equality}}

\begin{minted}[xleftmargin=1em,fontsize=\footnotesize]{haskell}
Get xs s == Just (ListOf x) || Not (Member xs s)
\end{minted}

To avoid adressing the type of value (as it may not exist at all), we defined an
 auxiliary predicate \mintinline{haskell}{IsList :: Maybe * -> Bool} to replace
 the former part.

\begin{minted}[xleftmargin=1em,fontsize=\footnotesize]{haskell}
IsList (Get xs s) || Not (Member xs s)
\end{minted}

The type expression above has kind \mintinline{haskell}{Bool}, we could make it
 a type constraint by asserting equality.

\begin{minted}[xleftmargin=1em,fontsize=\footnotesize]{haskell}
(IsList (Get xs s) || Not (Member xs s)) ~ True
\end{minted}

With \emph{constraint kind}, a recent addition to GHC. Type constraints now has
its own kind: \mintinline{haskell}{Constraint}. That means type constraints
are not restricted to the left side of a \mintinline{haskell}{=>} anymore,
they could appear in anywhere that accepts something of kind
\mintinline{haskell}{Constraint}, and any type that has kind
\mintinline{haskell}{Constraint} can also be used as a type constraint.\footnote{See \url{https://downloads.haskell.org/~ghc/7.4.1/docs/html/users_guide/constraint-kind.html}.}
We could give the lengthy constraint above a type synonym.

\begin{minted}[xleftmargin=1em,fontsize=\footnotesize]{haskell}
ListOrNX xs s =
    (IsList (Get xs s) || Not (Member xs s)) ~ True
\end{minted}

\section{Discussions}

\section{Conclusions and Related Work}
